---
title: "star protocol roi data from 3.6 segmentation"
output: html_document
date: "2024-11-19"
---

```{r setup, include=FALSE}
library(AbiMatrix)
library(Cardinal)
library(dplyr)
library(EBImage)
library(ggplot2)
library(reshape2)
library(tidyr)
```

Get ROI

```{r}
#setwd("~/Library/CloudStorage/OneDrive-TheMountSinaiHospital/Watit - BigPC/mouse-histology/control_cv")
c5_brain <- getSegmentROI(c5_ssc, cluster = 4, select_classes = c(1:3,5:13,15:20))
c5_cortex <- getSegmentROI(c5_ssc, cluster = 5, select_classes = c(6,8), select_target = 1)
c5_hippo <- getSegmentROI(c5_ssc, cluster = 5, select_classes = c(7,8,11), select_target = 8)
c5_cereb <- getSegmentROI(c5_ssc, cluster = 4, select_classes = c(2,5,15,16,17,19), select_target = 2)
```


```{r}
#setwd("~/Library/CloudStorage/OneDrive-TheMountSinaiHospital/Watit - BigPC/mouse-histology/control_cv")
c9_brain <- getSegmentROI(c9_ssc, cluster = 1, select_classes = c(1:8,10,12:19))
c9_cortex <- fillHull(getSegmentROI(c9_ssc, cluster = 1, select_classes = c(7,5), select_target = 1) +
  getSegmentROI(c9_ssc, cluster = 1, select_classes = c(7), select_target = 3) +
  getSegmentROI(c9_ssc, cluster = 1, select_classes = c(6), select_target = 2) +
  getSegmentROI(c9_ssc, cluster = 1, select_classes = c(8), select_target = 2) +
  getSegmentROI(c9_ssc, cluster = 1, select_classes = c(12), select_target = 1))
c9_cortex <- c9_cortex > 0
c9_hippo <- getSegmentROI(c9_ssc, cluster = 4, select_classes = c(6,10,15), select_target = 5)
c9_cereb <- getSegmentROI(c9_ssc, cluster = 1, select_classes = c(13,14), select_target = 1)
```


```{r}
#setwd("~/Library/CloudStorage/OneDrive-TheMountSinaiHospital/Watit - BigPC/mouse-histology/control_cv")
c10_brain <- getSegmentROI(c10_ssc, cluster = 3, select_classes = c(2:5, 7:13,15:20))
c10_cortex <- getSegmentROI(c10_ssc, cluster = 4, select_classes = c(5,12), select_target = 1)
c10_hippo <- getSegmentROI(c10_ssc, cluster = 4, select_classes = c(2,5,10,12), select_target = 13)
c10_cereb <- fillHull(getSegmentROI(c10_ssc, cluster = 4, select_classes = c(10,13,15,18), select_target = 1) +
  getSegmentROI(c10_ssc, cluster = 4, select_classes = c(10,13,15,18), select_target = 2) +
  getSegmentROI(c10_ssc, cluster = 4, select_classes = c(10,13,15,18), select_target = 3) +
  getSegmentROI(c10_ssc, cluster = 4, select_classes = c(9), select_target = 2))
c10_cereb <- c10_cereb > 0
```

resize image

```{r}

matchSize <- function(img, target) {
  
  w <- max(target@elementMetadata$x)
  h <- max(target@elementMetadata$y)
  
  resized_img <- resize(img, w, h)
  
  if (is.logical(img)) {
    resized_img <- resized_img > 0
  }
  
  return(resized_img)
}

c5_cereb <- matchSize(c5_cereb, c5_mse_proc)
c5_cortex <- matchSize(c5_cortex, c5_mse_proc)
c5_hippo <- matchSize(c5_hippo, c5_mse_proc)

c9_cereb <- matchSize(c9_cereb, c9_mse_proc)
c9_cortex <- matchSize(c9_cortex, c9_mse_proc)
c9_hippo <- matchSize(c9_hippo, c9_mse_proc)

c10_cereb <- matchSize(c10_cereb, c10_mse_proc)
c10_cortex <- matchSize(c10_cortex, c10_mse_proc)
c10_hippo <- matchSize(c10_hippo, c10_mse_proc)
```

Add ROIs to MSI objects

```{r}
c5_rois <- c("c5_cortex", "c5_hippo", "c5_cereb")
c5_roi_list <- mget(c5_rois)
c5_mse_proc <- assignROI(c5_mse_proc, c5_roi_list)

c9_rois <- c("c9_cortex", "c9_hippo", "c9_cereb")
c9_roi_list <- mget(c9_rois)
c9_mse_proc <- assignROI(c9_mse_proc, c9_roi_list)

c10_rois <- c("c10_cortex", "c10_hippo", "c10_cereb")
c10_roi_list <- mget(c10_rois)
c10_mse_proc <- assignROI(c10_mse_proc, c10_roi_list)
```

Check data

```{r}
# Alanine
c5_ala_msi <- image(c5_mse_proc, mz = c(90.055509, 92.071169), subset=c5_mse_proc$c5_cereb, enhance="adaptive", scale=TRUE)
# GABA
c5_gaba_msi <- image(c5_mse_proc, mz = c(104.071159, 106.086819), subset=c5_mse_proc$c5_cereb, enhance="adaptive", scale=TRUE)
# Aspartate
c5_asp_msi <- image(c5_mse_proc, mz = c(134.045339, 136.045999), subset=c5_mse_proc$c5_cereb, enhance="adaptive", scale=TRUE)
# Glutamate
c5_glu_msi <- image(c5_mse_proc, mz = c(147.076973, 149.082633), subset=c5_mse_proc$c5_cereb, enhance="adaptive", scale=TRUE)
#c5_glu_msi <- image(c5_mse_proc, mz = c(147.076973, 148.074803,	149.082633,	150.090463,	151.118293,	152.106123), subset=c5_mse_proc$c5_cereb, enhance="adaptive", scale=TRUE)
# Glutamine
c5_gln_msi <- image(c5_mse_proc, mz = c(148.060989, 150.06649), subset=c5_mse_proc$c5_cereb, enhance="adaptive", scale=TRUE)
#c5_gln_msi <- image(c5_mse_proc, mz = c(148.060989,	149.068819,	150.066649,	151.094479,	152.102309,	153.100139), subset=c5_mse_proc$c5_cereb, enhance="adaptive", scale=TRUE)

# GABA - c9
c9_gaba_msi <- image(c9_mse_proc, mz = c(104.071159, 106.086819), subset=c9_mse_proc$c9_cereb, enhance="adaptive", scale=TRUE)
# Aspartate - c9
c9_asp_msi <- image(c9_mse_proc, mz = c(134.045339, 136.045999), subset=c9_mse_proc$c9_cereb, enhance="adaptive", scale=TRUE)
# Glutamate - c9
c9_glu_msi <- image(c9_mse_proc, mz = c(147.076973, 149.082633), subset=c9_mse_proc$c9_cereb, enhance="adaptive", scale=TRUE)
#c9_glu_msi <- image(c9_mse_proc, mz = c(147.076973, 148.074803,	149.082633,	150.090463,	151.118293,	152.106123), subset=c9_mse_proc$c9_cereb, enhance="adaptive", scale=TRUE)
# Glutamine - c9
c9_gln_msi <- image(c9_mse_proc, mz = c(148.060989, 150.06649), subset=c9_mse_proc$c9_cereb, enhance="adaptive", scale=TRUE)
#c9_gln_msi <- image(c9_mse_proc, mz = c(148.060989,	149.068819,	150.066649,	151.094479,	152.102309,	153.100139), subset=c9_mse_proc$c9_cereb, enhance="adaptive", scale=TRUE)

# GABA - c10
c10_gaba_msi <- image(c10_mse_proc, mz = c(104.071159, 106.086819), subset=c10_mse_proc$c10_cereb, enhance="adaptive", scale=TRUE)
# Aspartate - c10
c10_asp_msi <- image(c10_mse_proc, mz = c(134.045339, 136.045999), subset=c10_mse_proc$c10_cereb, enhance="adaptive", scale=TRUE)
# Glutamate - c10
c10_glu_msi <- image(c10_mse_proc, mz = c(147.076973, 149.082633), subset=c10_mse_proc$c10_cereb, enhance="adaptive", scale=TRUE)
#c10_glu_msi <- image(c10_mse_proc, mz = c(147.076973, 148.074803,	149.082633,	150.090463,	151.118293,	152.106123), subset=c10_mse_proc$c10_cereb, enhance="adaptive", scale=TRUE)
# Glutamine - c10
c10_gln_msi <- image(c10_mse_proc, mz = c(148.060989, 150.06649), subset=c10_mse_proc$c10_cereb, enhance="adaptive", scale=TRUE)
#c10_gln_msi <- image(c10_mse_proc, mz = c(148.060989,	149.068819,	150.066649,	151.094479,	152.102309,	153.100139), subset=c10_mse_proc$c10_cereb, enhance="adaptive", scale=TRUE)
```

Create function to extract the tic

```{r}
getROITIC <- function(msi, roi_list, normalize = TRUE, decimal = 4, verbose = FALSE) {
  df <- NULL
  for (name in names(roi_list)) {
    if (verbose) {
      cat("Get TIC from ROI:", name, "\n")
    }
    tic <- fData(summarizeFeatures(msi, stat="sum", groups=msi[[name]], verbose = FALSE))
    if (is.null(df)) {
      df <- data.frame(mz = as.numeric(format(round(as.matrix(tic[,"mz"]), decimal), nsmall = decimal)))
    }

    if (normalize) {
      new_df <- data.frame(tic = as.matrix(tic[,"TRUE.sum"])/sum(roi_list[[name]]))

    } else {
      new_df <- data.frame(tic = as.matrix(tic[,"TRUE.sum"]))
    }

    colnames(new_df) <- name
    df <- cbind(df, new_df)
  }
  return(df)
}

```

Extract TIC from MSI objects

```{r}
c5_data <- getROITIC(c5_mse_proc, c5_roi_list, normalize = FALSE)
c9_data <- getROITIC(c9_mse_proc, c9_roi_list, normalize = FALSE)
c10_data <- getROITIC(c10_mse_proc, c10_roi_list, normalize = FALSE)

c5_ndata <- getROITIC(c5_mse_proc, c5_roi_list, normalize = TRUE)
c9_ndata <- getROITIC(c9_mse_proc, c9_roi_list, normalize = TRUE)
c10_ndata <- getROITIC(c10_mse_proc, c10_roi_list, normalize = TRUE)
```

Create data frame for reproducibility

```{r}
tic_list <- mget(c("c5_data", "c9_data", "c10_data"))
col_list <- c("Cortex", "Hippocampus", "Cerebellum")
cv_df <- buildCVDataframe(tic_list, col_list)
cv <- getCV(cv_df, col_list, na.rm = TRUE, outline.rm = TRUE, noise.rm = TRUE)

ggplot(cv, aes(x = Region, y = CV, fill = Region)) +
  geom_errorbar(aes(ymin = CV - 3*SD, ymax = CV + 3*SD), width = 0.2) +
  geom_boxplot(aes(lower=CV-SD, upper=CV+SD, middle=CV, ymin=CV-3*SD, ymax=CV+3*SD), 
               stat = "identity",) +
  labs(title = "Coefficient of Variation by Region",
       x = "Region",
       y = "CV (%)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5))
```

Create some function for filtering data

```{r}
filterMetaboliteTIC <- function(msi_df, ion_list, tracing=NULL) {
  msi_tic <- list()
  if (is.data.frame(ion_list)) {
    metabolite_names <- ion_list[, unlist(lapply(ion_list, is.character), use.names = FALSE)]
    mass_list <- ion_list[, unlist(lapply(ion_list, is.numeric), use.names = FALSE)]
    if (!is.null(tracing)) {
      new_names <- NULL
      new_masses <- NULL
      new_df <- NULL
      for (i in tracing) {
        new_names <- c(new_names, paste0(metabolite_names,"_M",i))
        new_masses <- c(new_masses, mass_list + (i*1.00783))
      }
      new_df <- data.frame(Metabolite=new_names, mz=new_masses)
      new_df[order(new_df$Metabolite),]
      metabolite_names <- new_df$Metabolite
      mass_list <- new_df$mz
    }
    
    head_col <- data.frame(Metabolite=metabolite_names, mz=mass_list)
  } else {
    mass_list <- ion_list
    head_col <- data.frame(mz=mass_list)
  }
  
  for (i in seq_along(mass_list)) {
    mass <- mass_list[i]
    row_number <- which.min(abs(msi_df$mz - mass))
    col_number <- which(colnames(msi_df)=="mz") + 1
    new_tic <- cbind(head_col[i, , drop=FALSE], msi_df[row_number, col_number:ncol(msi_df)])
    msi_tic <- rbind(msi_tic, new_tic)
  }

  return(msi_tic)
}

```

Compile in one dataframe for selected compounds

```{r}
tic_list <- mget(c("c5_data", "c9_data", "c10_data"))
col_list <- c("Cortex", "Hippocampus", "Cerebellum")
customed_metabolites <- data.frame(
  Metabolite = c("GABA_M0", "GABA_M2", "Aspartate_M0", "Aspartate_M2",  "Glutamate_M0", "Glutamate_M2", "Glutamine_M0", "Glutamine_M2"),
  Mass = c(104.071159, 106.086819, 134.045339, 136.060999, 147.076973, 149.082633, 148.060989, 150.06649)
)

c5_filtered <- filterMetaboliteTIC(c5_data, customed_metabolites)
c9_filtered <- filterMetaboliteTIC(c9_data, customed_metabolites)
c10_filtered <- filterMetaboliteTIC(c10_data, customed_metabolites)

merged_df <- merge(merge(c5_filtered, c9_filtered), c10_filtered)

c5_nfiltered <- filterMetaboliteTIC(c5_ndata, customed_metabolites)
c9_nfiltered <- filterMetaboliteTIC(c9_ndata, customed_metabolites)
c10_nfiltered <- filterMetaboliteTIC(c10_ndata, customed_metabolites)

merged_ndf <- merge(merge(c5_nfiltered, c9_nfiltered), c10_nfiltered)
```

Plot

```{r}
# filter GABA data
gaba_data <- merged_df[grep("GABA", merged_df$Metabolite), ]

# reshape the data for easier plotting
gaba_long <- melt(
  gaba_data,
  id.vars = c("Metabolite", "mz"),
  variable.name = "Region",
  value.name = "Intensity"
)

# extract Region and Sample details from column names
gaba_long <- gaba_long %>%
  mutate(
    Metabolite = sub(".*_(M[0-9]+)", "\\1", Metabolite),
    Sample = sub("_.*", "", Region),        # extract sample name (e.g., "c5", "c9", "c10")
    Region = gsub("c[0-9]+_", "", Region)  # remove sample prefixes to get region names
  )

# compute summary statistics (mean and standard error) for each group
summary_data <- gaba_long %>%
  group_by(Metabolite, Region) %>%
  summarize(
    Mean = mean(Intensity, na.rm = TRUE),
    SE = sd(Intensity, na.rm = TRUE) / sqrt(n()),  # Standard error
    .groups = "drop"
  )

# create the plot
ggplot(summary_data, aes(x = Region, y = Mean, fill = Metabolite)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black", width = 0.7) +  # Bar plot
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE), position = position_dodge(width = 0.9), width = 0.2) +  # Error bars
  geom_jitter(data = gaba_long, aes(x = Region, y = Intensity, color = Metabolite), 
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.9), size = 2, alpha = 0.8) +  # Dots for individual points
  scale_fill_brewer(palette = "Set1") +  # customize bar colors
  scale_color_brewer(palette = "Set1") +  # match dot colors to bar colors
  theme_minimal() +
  labs(
    title = "GABA Intensities Across Regions",
    x = "Brain Region",
    y = "Normalized Intensity",
    fill = "Isotopologues",
    color = "Isotopologues"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_text(size = 16),  # Increase x-axis title size
        axis.title.y = element_text(size = 16),  # Increase y-axis title size
        plot.title = element_text(size = 18, hjust = 0.5)  # Increase title size and center it
        )
```

Let create function and applied to all other metabolites

```{r}
plotMetabolite <- function(data, metabolite_name) {
  # filter data for the specified metabolite
  metabolite_data <- data[grep(metabolite_name, data$Metabolite), ]
  
  # reshape the data for easier plotting
  metabolite_long <- melt(
    metabolite_data,
    id.vars = c("Metabolite", "mz"),
    variable.name = "Region",
    value.name = "Intensity"
  )
  
  # extract Region and Sample details from column names
  metabolite_long <- metabolite_long |>
    mutate(
      Metabolite = sub(".*_(M[0-9]+)", "\\1", Metabolite),
      Sample = sub("_.*", "", Region),        # Extract sample name (e.g., "c5", "c9", "c10")
      Region = gsub("c[0-9]+_", "", Region)  # Remove sample prefixes to get region names
    )
  
  # compute summary statistics (mean and standard error) for each group
  summary_data <- metabolite_long |>
    group_by(Metabolite, Region) |>
    summarize(
      Mean = mean(Intensity, na.rm = TRUE),
      SE = sd(Intensity, na.rm = TRUE) / sqrt(n()),  # Standard error
      .groups = "drop"
    )
  
  ggplot(summary_data, aes(x = Region, y = Mean, fill = Metabolite)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black", width = 0.7) +  # Bar plot
    geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE), position = position_dodge(width = 0.9), width = 0.2) +  # Error bars
    geom_jitter(data = metabolite_long, aes(x = Region, y = Intensity, fill = Metabolite), 
                position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.9), shape = 21, size = 2, color = "black", alpha = 0.8) +  # Dots for individual points
    scale_fill_brewer(palette = "Set1") +  # Customize bar colors
    scale_color_brewer(palette = "Set1") +  # Match dot colors to bar colors
    theme_minimal() +
    labs(
      title = paste(metabolite_name, "Abundances Across Regions"),
      x = "Brain Region",
      y = paste0(metabolite_name," Intensities"),
      fill = "Isotopologue",
      color = "Isotopologue"
    ) +
    theme(
      axis.text.x = element_text(size = 16, angle = 45, hjust = 1),  # Increase size of x-axis labels
      axis.title.x = element_text(size = 18),  # Increase size of x-axis title
      axis.title.y = element_text(size = 18),  # Increase size of y-axis title
      plot.title = element_text(size = 20, hjust = 0.5)  # Increase size of title and center it
    )
}

# Plot examples
p_gaba <- plotMetabolite(merged_df, "GABA")
p_asp <- plotMetabolite(merged_df, "Aspartate")
p_glu <- plotMetabolite(merged_df, "Glutamate")
p_gln <- plotMetabolite(merged_df, "Glutamine")

p_ngaba <- plotMetabolite(merged_ndf, "GABA")
p_nasp <- plotMetabolite(merged_ndf, "Aspartate")
p_nglu <- plotMetabolite(merged_ndf, "Glutamate")
p_ngln <- plotMetabolite(merged_ndf, "Glutamine")
```

Create new enrichment dataframe

```{r}
normalizeTracer <- function(data) {
  # Step 1: Calculate totals for each base metabolite
  totals_df <- data %>%
    mutate(BaseMetabolite = gsub("_M[0-9]+$", "", Metabolite)) %>%
    group_by(BaseMetabolite) %>%
    summarize(across(where(is.numeric), sum), .groups = "drop") %>%
    rename_with(~ paste0("Total_", .), -BaseMetabolite)  # Prefix totals with "Total_"

  # Step 2: Join the totals back to the original dataframe
  data_with_proportions <- data %>%
    mutate(BaseMetabolite = gsub("_M[0-9]+$", "", Metabolite)) %>%
    left_join(totals_df, by = "BaseMetabolite") %>%
    # Step 3: Calculate proportions for M0 and M2
    mutate(across(
      where(is.numeric) & !starts_with("Total_"),
      ~ .x / get(paste0("Total_", cur_column())),
      .names = "Proportion_{col}"
    )) %>%
    # Step 4: Filter to keep only M0 and M2 rows and proportion columns
    filter(grepl("_M[0-9]+$", Metabolite)) %>%
    select(Metabolite, mz, starts_with("Proportion_"))

  # Step 5: Clean up column names (remove Proportion_ prefix and drop Proportion_mz)
  final_df <- data_with_proportions %>%
    select(-Proportion_mz) %>%
    rename_with(~ gsub("Proportion_", "", .), starts_with("Proportion_"))
  
  return(final_df)
}

merged_edf <- normalizeTracer(merged_df)
p_egaba <- plotMetabolite(merged_edf, "GABA")
p_easp <- plotMetabolite(merged_edf, "Aspartate")
p_eglu <- plotMetabolite(merged_edf, "Glutamate")
p_egln <- plotMetabolite(merged_edf, "Glutamine")

merged_endf <- normalizeTracer(merged_ndf)
p_engaba <- plotMetabolite(merged_endf, "GABA")
p_enasp <- plotMetabolite(merged_endf, "Aspartate")
p_englu <- plotMetabolite(merged_endf, "Glutamate")
p_engln <- plotMetabolite(merged_endf, "Glutamine")
```

Plot compile

```{r}
library(patchwork)

p <- (p_glu + p_gaba) / (p_gln + p_asp)
p_n <- (p_nglu + p_ngaba) / (p_ngln + p_nasp)
p_e <- (p_eglu + p_egaba) / (p_egln + p_easp)
p_en <- (p_englu + p_engaba) / (p_engln + p_enasp)
```

```{r}
library(ggpubr)

plotMetabolite2 <- function(data, metabolite_name) {
  # Filter data for the specified metabolite
  metabolite_data <- data[grep(metabolite_name, data$Metabolite), ]
  
  # Reshape the data for easier plotting
  metabolite_long <- melt(
    metabolite_data,
    id.vars = c("Metabolite", "mz"),
    variable.name = "Region",
    value.name = "Intensity"
  )
  
  # Extract Region and Sample details from column names
  metabolite_long <- metabolite_long |>
    mutate(
      Metabolite = sub(".*_(M[0-9]+)", "\\1", Metabolite),
      Sample = sub("_.*", "", Region),        # Extract sample name (e.g., "c5", "c9", "c10")
      Region = gsub("c[0-9]+_", "", Region)  # Remove sample prefixes to get region names
    )
  
  # Compute summary statistics (mean and standard error) for each group
  summary_data <- metabolite_long |>
    group_by(Metabolite, Region) |>
    summarize(
      Mean = mean(Intensity, na.rm = TRUE),
      SE = sd(Intensity, na.rm = TRUE) / sqrt(n()),  # Standard error
      .groups = "drop"
    )
  
  # Conduct pairwise t-tests between regions for each Metabolite
  stat_test <- metabolite_long %>%
    group_by(Metabolite) %>%
    t_test(Intensity ~ Region, paired = FALSE) %>%
    adjust_pvalue(method = "bonferroni") %>%
    mutate(
      p.adj.signif = case_when(
        p.adj <= 0.001 ~ "***",
        p.adj <= 0.01 ~ "**",
        p.adj <= 0.05 ~ "*",
        TRUE ~ "NS"
      )
    )
  
  # Add position for significance annotations
  stat_test <- stat_test %>%
    mutate(y.position = max(summary_data$Mean) + SE * 1.5)

  # Plot with p-value annotations
  ggplot(summary_data, aes(x = Region, y = Mean, fill = Metabolite)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black", width = 0.7) +  # Bar plot
    geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE), position = position_dodge(width = 0.9), width = 0.2) +  # Error bars
    geom_jitter(data = metabolite_long, aes(x = Region, y = Intensity, fill = Metabolite), 
                position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.9), shape = 21, size = 2, color = "black", alpha = 0.8) +  # Dots for individual points
    scale_fill_brewer(palette = "Set1") +  # Customize bar colors
    scale_color_brewer(palette = "Set1") +  # Match dot colors to bar colors
    theme_minimal() +
    labs(
      title = paste(metabolite_name, "Abundances Across Regions"),
      x = "Brain Region",
      y = paste0(metabolite_name, " Intensities"),
      fill = "Isotopologue",
      color = "Isotopologue"
    ) +
    theme(
      axis.text.x = element_text(size = 16, angle = 45, hjust = 1),  # Increase size of x-axis labels
      axis.title.x = element_text(size = 18),  # Increase size of x-axis title
      axis.title.y = element_text(size = 18),  # Increase size of y-axis title
      plot.title = element_text(size = 20, hjust = 0.5)  # Increase size of title and center it
    ) +
    stat_pvalue_manual(stat_test, label = "p.adj.signif", tip.length = 0.01)
}

```


Extract new set of metabolites

```{r}
ala_isotopologue <- data.frame(
  Metabolite = c("Alanine_M0", "Alanine_M1", "Alanine_M2", "Alanine_M3"),
  Mass = c(90.055509,	91.063339,	92.071169,	93.078999)
)

gaba_isotopologue <- data.frame(
  Metabolite = c("GABA_M0", "GABA_M1", "GABA_M2", "GABA_M3",  "GABA_M4"),
  Mass = c(104.071159,	105.078989,	106.086819,	107.094649,	108.102479)
)

glu_isotopologue <- data.frame(
  Metabolite = c("Glutamate_M0", "Glutamate_M1", "Glutamate_M2", "Glutamate_M3",  "Glutamate_M4", "Glutamate_M5"),
  Mass = c(147.076973,	148.084803,	149.092633,	150.100463,	151.108293,	152.116123)
)

gln_isotopologue <- data.frame(
  Metabolite = c("Glutamine_M0", "Glutamine_M1", "Glutamine_M2", "Glutamine_M3",  "Glutamine_M4", "Glutamine_M5"),
  Mass = c(148.060989,	149.068819,	150.076649,	151.084479,	152.092309,	153.100139)
)

four_metabolites <- rbind(ala_isotopologue, gaba_isotopologue)
four_metabolites <- rbind(four_metabolites, glu_isotopologue)
four_metabolites <- rbind(four_metabolites, gln_isotopologue)

c5_nfiltered2 <- filterMetaboliteTIC(c5_ndata, four_metabolites)
c9_nfiltered2 <- filterMetaboliteTIC(c9_ndata, four_metabolites)
c10_nfiltered2 <- filterMetaboliteTIC(c10_ndata, four_metabolites)

merged_ndf2 <- merge(merge(c5_nfiltered2, c9_nfiltered2), c10_nfiltered2)

```


```{r}
# Use Amir peaks H = 1.00331
ala_isotopologue <- data.frame(
  Metabolite = c("Alanine_M0", "Alanine_M1", "Alanine_M2", "Alanine_M3"),
  Mass = c(90.055479,	91.058779,	92.062079,	93.065379)
)

asp_isotopologue <- data.frame(
  Metabolite = c("Aspartate_M0", "Aspartate_M1", "Aspartate_M2", "Aspartate_M3", "Aspartate_M4"),
  Mass = c(134.045259,	135.048509,	136.051759,	137.055009,	138.058259)
)

gaba_isotopologue <- data.frame(
  Metabolite = c("GABA_M0", "GABA_M1", "GABA_M2", "GABA_M3",  "GABA_M4"),
  Mass = c(104.07095,	105.07434,	106.07773,	107.08112,	108.08451)
)

glu_isotopologue <- data.frame(
  Metabolite = c("Glutamate_M0", "Glutamate_M1", "Glutamate_M2", "Glutamate_M3",  "Glutamate_M4", "Glutamate_M5"),
  Mass = c(148.06102,	149.06433,	150.06764,	151.07095,	152.07426,	153.07757)
)

gln_isotopologue <- data.frame(
  Metabolite = c("Glutamine_M0", "Glutamine_M1", "Glutamine_M2", "Glutamine_M3",  "Glutamine_M4", "Glutamine_M5"),
  Mass = c(147.07694,	148.060065,	149.08319,	150.072315,	151.06944,	152.042565)
)



four_metabolites <- NULL
four_metabolites <- rbind(four_metabolites, ala_isotopologue)
four_metabolites <- rbind(four_metabolites, asp_isotopologue)
four_metabolites <- rbind(four_metabolites, gaba_isotopologue)
four_metabolites <- rbind(four_metabolites, glu_isotopologue)
four_metabolites <- rbind(four_metabolites, gln_isotopologue)

c5_nfiltered3 <- filterMetaboliteTIC(c5_ndata, four_metabolites)
c9_nfiltered3 <- filterMetaboliteTIC(c9_ndata, four_metabolites)
c10_nfiltered3 <- filterMetaboliteTIC(c10_ndata, four_metabolites)

merged_ndf3 <- merge(merge(c5_nfiltered3, c9_nfiltered3), c10_nfiltered3)

p_agaba <- plotMetabolite(merged_ndf3, "GABA")
p_aala <- plotMetabolite(merged_ndf3, "Alanine")
p_aasp <- plotMetabolite(merged_ndf3, "Aspartate")
p_aglu <- plotMetabolite(merged_ndf3, "Glutamate")
p_agln <- plotMetabolite(merged_ndf3, "Glutamine")

p_a <- (p_aglu + p_agaba) / (p_agln + p_aasp) / p_aala

```


```{r}
pv_isotopologue <- data.frame(
  Metabolite = c("Pyruvate_M0", "Pyruvate_M1", "Pyruvate_M2", "Pyruvate_M3"),
  Mass = c(89.053845,	90.031645,	91.039445,	92.047245)
)

gaba_isotopologue <- data.frame(
  Metabolite = c("GABA_M0", "GABA_M1", "GABA_M2", "GABA_M3",  "GABA_M4"),
  Mass = c(104.07095,	105.07434,	106.07773,	107.08112,	108.08451)
)

sc_isotopologue <- data.frame(
  Metabolite = c("Succinate_M0", "Succinate_M1", "Succinate_M2", "Succinate_M3", "Succinate_M4"),
  Mass = c(119.08481,	120.1221,	121.12401,	122.08781,	123.07561)
)

fm_isotopologue <- data.frame(
  Metabolite = c("Fumarate_M0", "Fumarate_M1", "Fumarate_M2", "Fumarate_M3", "Fumarate_M4"),
  Mass = c(117.05876,	118.02656,	119.04436,	120.04216,	121.04996)
)

gln_isotopologue <- data.frame(
  Metabolite = c("Glutamine_M0", "Glutamine_M1", "Glutamine_M2", "Glutamine_M3",  "Glutamine_M4", "Glutamine_M5"),
  Mass = c(147.07694,	148.060065,	149.08319,	150.072315,	151.06944,	152.042565)
)

glu_isotopologue <- data.frame(
  Metabolite = c("Glutamate_M0", "Glutamate_M1", "Glutamate_M2", "Glutamate_M3",  "Glutamate_M4", "Glutamate_M5"),
  Mass = c(148.06102,	149.06433,	150.06764,	151.07095,	152.07426,	153.07757)
)


six_metabolites <- NULL
six_metabolites <- rbind(six_metabolites, pv_isotopologue)
six_metabolites <- rbind(six_metabolites, gaba_isotopologue)
six_metabolites <- rbind(six_metabolites, sc_isotopologue)
six_metabolites <- rbind(six_metabolites, fm_isotopologue)
six_metabolites <- rbind(six_metabolites, gln_isotopologue)
six_metabolites <- rbind(six_metabolites, glu_isotopologue)

c5_6mets <- filterMetaboliteTIC(c5_ndata, six_metabolites)
c9_6mets <- filterMetaboliteTIC(c9_ndata, six_metabolites)
c10_6mets <- filterMetaboliteTIC(c10_ndata, six_metabolites)

merged_final <- merge(merge(c5_6mets, c9_6mets), c10_6mets)

p_pv <- plotMetabolite(merged_final, "Pyruvate")
p_gaba <- plotMetabolite(merged_final, "GABA")
p_sc <- plotMetabolite(merged_final, "Succinate")
p_fm <- plotMetabolite(merged_final, "Fumarate")
p_gln <- plotMetabolite(merged_final, "Glutamine")
p_glu <- plotMetabolite(merged_final, "Glutamate")

p_a <- (p_pv + p_gaba) / (p_sc + p_fm) / (p_gln + p_glu)

write.csv(merged_final, "final_msi_data.csv", row.names = FALSE)

# code location on labarchive: https://mynotebook.labarchives.com/NzYyOTM4Ljh8NTg2ODc2LzU4Njg3Ni9Ob3RlYm9vay8xMjgwOTM4ODM1fDE5MzY2OTAuNzk5OTk5OTk5OA==/page/1712891-2606
```